---
title: "PS270 Final Project"
author: Hannah Paley 
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction 

  Does personally knowing someone who has had an abortion impact an individual's
stance on abortion policy? Within this study I am looking at whether knowing 
or not knowing someone who has had an abortion impacts how an individual views
abortion legislation. I hypothesize that personally knowing someone who has
had an abortion will positively correlate to somewhat or strongly opposing a 
proposed 16-week abortion ban. I believe that being personally connected to 
someone who has had an abortion makes it less likely that an individual will
support legislation that decreases abortion access. Having this personal 
connection to abortion may make it so people feel the issue is "closer" 
to them than those who do not have a personal connection to abortion. 
This could help future public policy as well as elected officials figure out
how to campaign. Overall it has the ability to help better understand why
people think the way they do about an issue as polarizing as abortion in the 
United States. 

  My data sample is based off of a web-survey measuring individual abortion 
attitudes between February 23, 2024 - February 25, 2024. My unit of analysis 
is an individual person. The explanatory variable of interest is whether or
not a surveyed individual knows someone who has had an abortion. The variable
is initially coded as yes, yes, I have, or yes I know someone if the 
individual knows someone who has had an abortion, or no if the individual 
does not someone who has had an abortion. My outcome variable is whether an 
individual strongly supports, somewhat supports, somewhat opposes, or strongly
opposes a nationwide ban on abortion at 16 weeks of pregnancy which I will 
eventually turn into a 4-point scale for a linear regression. If I observe an 
increase in "strongly opposes" and "somewhat opposes" than other responses 
regarding my outcome variable (going in a more negative direction), this 
would support my hypothesis when looking at the shift in not knowing to 
knowing someone who has had an abortion. Conversely, if I observe an increase
in "strongly supports" and "somewhat supports" (going in a more positive
direction) than other responses regarding my outcome variable, this would go
against my hypothesis. A negative correlation between the shift from not
knowing to knowing someone who has had an abortion and attitudes surrounding
a 16-week abortion ban (with 1 as strongly opposed and 4 as strongly 
supported, therefore a decrease in this attitude would point to a negative 
stance on restrictive abortion policies) would indicate support for 
my hypothesis. 

## Data Section 

  The provided data is measured by a web-based survey produced by The Roper 
Center. There were 1020 respondents and 42 different variables. The independent
variable in this specific project is whether an individuals knows someone who 
has had an abortion. The dependent variable is their attitude towards a 
proposed 16-week abortion ban.

  My research design is a cross-sectional design, as I am assessing the data at 
one point in time where no variable is manipulated. This assumes that the 
groups are identical on average (those who know someone and those who don't 
know someone who has had an abortion).

```{r Dependent-Variable}
library(tidyverse)
abortions <- read_csv("Data/Final Project Data.csv")

a_attitudes <- abortions |>
  group_by(...13) |> summarize(n = n()) 
a_attitudes
a_attitudes <- a_attitudes |> 
filter(...13 != "BNE2_2") |> filter(...13 != "Skipped")
ggplot(data = a_attitudes, 
       mapping = aes(x = ...13, y = n)) + 
  geom_col(fill = "skyblue3") +
  theme(panel.background = element_blank()) + 
  labs(x = "Response to Sixteen Week Abortion Ban", 
       y = "Number of Responses", title = "Legislation Response",
       caption = 
         "Roper Abortion Attitudes / Alabama Supreme Court") 
```

## Results 
```{r Connected-Abortion}

abortions_yes <-
  abortions |> 
  mutate( yes = ...17 == "Yes" | 
                         ...17 == "Yes, I have"| 
          ...17 == "Yes, I know someone") |>
  filter(yes)

abortions_yes |> select(...17, ...13) |> filter(...13 != "Skipped") |> 
  group_by(...13) |>
  summarize(n = n()) |>
  mutate(prop = n/sum(n)) |> 
  knitr::kable(col.names = c("Abortion Ban at 16 Weeks", 
      "Number of Answers", 
      "Prop. of Answers "), 
caption = "Responses to Abortion Ban by 
Those Connected to Abortion")

```
```{r Not-Connected-Abortion}
abortions_no <- abortions |> filter(...17 == "No")
abortions_no
abortions_no |> select(...17, ...13) |> filter(...13 != "Skipped") |>
  group_by(...13) |>
  summarize(n = n()) |>
  mutate(prop = n/sum(n)) |> 
  knitr::kable(col.names = c("Abortion Ban at 16 Weeks",
         "Number of Answers", 
         "Prop. of Answers "), 
      caption = "Responses to Abortion Ban by Those Not Connected to Abortion")
```

  Looking at both of the above tables, there seems to be a difference in the 
proportion of individuals who strongly oppose and somewhat oppose a 16-week 
abortion ban between those who know someone who has had an abortion and those
who don't. There is a higher proportion of individuals who strongly oppose and 
somewhat oppose an abortion ban among those who know someone who has had an 
abortion in comparison to individuals who do not. This points to a possible 
correlation between knowing someone who has had an abortion and being against 
an abortion ban. I will look into this further by running a linear regression 
and recoding the variables of interest.

```{r Impact-Abortion-Visual}
abortionsp <- abortions |> 
  mutate(...17new = if_else(...17=="Yes" 
                            | ...17 == "Yes, I have"
                            | ...17 == "Yes, I know someone", 1, 0)) |> 
  filter(...13 != "BNE2_2") |> 
  filter(...13 != "Skipped") |>  
  mutate(...13new = case_when(...13 == "Strongly oppose" ~ 1,
                              ...13 == "Somewhat oppose"~ 2,
                              ...13 == "Somewhat support" ~ 3, ...13 == "Strongly support" ~ 4)) 

ggplot(abortionsp, aes(x = ...17new, 
                       y = ...13new)) + 
  geom_point(colour = "skyblue3", alpha = 0) + stat_smooth(method = lm, colour = "slateblue1") + 
  labs( x = "Connection to Abortion", 
        y = "Legislation Response", 
        title = "Impact of Abortion on Legislation Preference"
        , caption = "Roper Abortion Attitudes / Alabama Supreme Court")
```

```{r Filter-Out}
library(infer)

abortions
abortionsp <-
  abortions |> filter(...17 != "BNE3_2") |> 
  filter(...17 != "Prefer not to answer") |> 
  filter(...17 != "Skipped")
```


```{r Re-Code}
abortionsp <- abortionsp |> 
  filter(...13 != "BNE2_2") |> 
  filter(...13 != "Skipped") |> 
  mutate(...17new = if_else(...17=="Yes" |
                              ...17 == "Yes, I have" | 
                              ...17 == "Yes, I know someone", 1, 0)) |>
  mutate(...13new = case_when(...13 == "Strongly oppose" ~ 1, ...13 == "Somewhat oppose"~ 2,
                              ...13 == "Somewhat support" ~ 3,
                              ...13 == "Strongly support" ~ 4))
```


```{r Regression-Output}
abortionsp |> select(...13new, ...17new)
abortion_ate <- lm(...13new ~ ...17new, data = abortionsp)  
abortion_ate
 var_labels <- c("(Intercept)" = "Intercept",
                  "...17new" = "Abortion Connection"
                  )
modelsummary::modelsummary(abortion_ate,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           coef_map = var_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared")) 
 


```
The Plot: The Impact of Abortion plot shows that as connection to abortion is 
increased from 0 to 1 on the x axis, and therefore we are shifting from those 
who do not know anyone who has had an abortion to those who do know someone who 
has had an abortion, there is a decrease in legislation response. What this
decrease represents is a more negative response to a 16-week abortion ban. 
There is an obvious negative slope that supports this statement.

Regression Output: The coefficent of Abortion Connection is  -.290. This can be
understood as meaning that attitudes about an abortion ban at 16 weeks get more 
negative and move in the direction of strongly oppose (1) as connection to
abortion increases from 0 to 1. As stated in my initial hypothesis, a negative 
coefficient shows that there is a negative correlation between knowing someone
who has had an abortion and supporting a 16-week abortion ban. This coefficient
is statistically significant, as the corresponding p-value is <0.001 indicating
high statistical significance. It would be dangerous to interpret this study 
causaully, as there could be various confounders that are not measured in this 
data set, so while we can interpret it as correlation, we should not interpret 
it as causation. To fully be confident that the coefficent represents a causal 
relationship, it would be imperative to run various permutations, statistical 
control, and the use of multiple subcalissifications. 
  
## Conclusion
  I found that there is a correlation between whether knowing or not knowing
someone who has had an abortion and how a proposed 16-week abortion ban is 
assessed, which mostly supports my hypothesis. As connection to abortion 
increases from 0 to 1, there is a decrease of about .29 in legislation response,
pointing to a decrease in positive attitudes towards an abortion ban. This 
decrease was statistically significant with a p-value < 0.001 proving that it is
fair to say there is a correlation between the independent and dependent 
variables. While there may be a correlation, I cannot confidently say there is 
causation, as there can be many things besides connection to abortion that could
cause an individual to respond negatively towards an abortion ban, such as the 
recent overturning of Roe v. Wade and how that is percieved. I think that 
although I talk about correlation in my hypothesis, I was also hypothesizing 
some type of causation by using the word "impacts" in my introduction which 
alludes to some type of causal effect. This part of my hypothesis is not 
supported by my data. 

  This points to one of the limitations of the analysis and threats to inference
which is confounding. This is especially present in my cross-sectional 
analysis, as confounding is one of the main issues within cross-sectional 
studies specifically. If I had more time/money, I think it would be interesting
to run permutations on my data and look at some of the other variables measured
in the study to strenghten the analysis. Additionally, I think it would have 
been interesting to further understand how connection to abortion impacts
response to other legislation besides just a 16-week abortion ban. 
  
## Work Cited

Axios. (2024). Axios/Ipsos Survey: Abortion Attitudes/Alabama Supreme Court 
(Version 2) [Dataset]. Roper Center for Public Opinion Research. doi:10.25940/
ROPER-31120887